---
title: "Practical ML Assignment"
author: "Sathya Prakash"
date: "Sunday, May 24, 2015"
output: html_document
---
Given Training Data is read using the using read.csv. It was observed that some of values are stored as NA string.

```{r}
setwd("D:\\R")
training<-read.csv("pml-training.csv",na.strings="")
training[training=="NA"]=NA
onlyNAs<-sapply(training,function(x) sum(is.na(x))>19000)
trainingDataClean<-training[!onlyNAs]
trainingDataClean<-trainingDataClean[sapply(trainingDataClean,function(x) is.numeric(x))]
trainingDataClean$classe<-training$classe

trainingDataClean<-subset(trainingDataClean,select=-c(X,raw_timestamp_part_1,raw_timestamp_part_2,num_window))
dim(trainingDataClean)
```

DO the same processing on the Test data as well

```{r}
testing<-read.csv("pml-testing.csv",na.strings="")
testing[testing=="NA"]=NA
testingDataClean<-testing[!onlyNAs]
testingDataClean<-testingDataClean[sapply(testingDataClean,function(x) is.numeric(x))]

testingDataClean<-subset(testingDataClean,select=-c(X,raw_timestamp_part_1,raw_timestamp_part_2,num_window))
dim(testingDataClean)
```
Observed that there are no covariates with near zero variance. With near Zero Variance such co variates will not be helpful in prediction.

```{r}
library(caret)
zeroVar = nearZeroVar(trainingDataClean,saveMetrics=TRUE)
any(zeroVar$nzv)

```

Split the Training Data into Training and Validation using CreateDataPartition
```{r}
inTrain<-createDataPartition(y=trainingDataClean$classe,p=0.7,list=FALSE)
t<-trainingDataClean[inTrain,]
v<-trainingDataClean[-inTrain,]
corr<-cor(t[,1:52])
library(corrplot)
corrplot(corr, order = "FPC", method = "color", type = "lower", tl.cex = 0.5, 
         tl.col = rgb(0, 0, 0))

```

Training Using Random Forests with cross validation of 5 times
```{r}
model<-train(t$classe~.,data=t,method="rf",trControl=trainControl(method="cv",5),ntree=251)
model
```
Out of Sample error is determined using the validation data selected earlier
```{r}
p<-predict(model,v)
acc<-postResample(p,v$classe)
acc
confusionMatrix(v$classe,p)
```

The Prediction on the test Data is obtained by applying the model
```{r}
p<-predict(model,testingDataClean)
p
```